A Guide to Designing Experiments to Test Statistical Graphics\thanks{The author(s) received no specific funding for this work.}

### 2025-06-04 index-revised.qmd
Revision updates

![Direct adjustment of a plot in a perceptual task. In this experiment, designed to assess the strength of the sine illusion, the user adjusts the plot using - and + buttons, which control the strength of a transformation designed to correct the effect of the sine illusion. When the user is satisfied that the lines are of equal length, they select the 'Finished' button to move to the next task. The experiment used a psychophysics experimental design, the method of adjustment, but leveraged the interactive Shiny interface to record the entire sequence of adjustments made by the user for each trial. A demo version of this application can be found at https://shiny.srvanderplas.com/sine-illusion/.  ](sine_illusion_screenshot.png){#fig-sine-illusion}

@@

![This question asks users to write out a description of how the population of Ewoks changes over time, without any further cues, to determine whether participants default to multiplicative or additive language descriptions.](Estimation_describe_plot_crop.png){#fig-estimation-describe}

@@

![This question asks participants for a numerical estimate, but provides a basic calculator and scratchpad. All user interactions with the calculator and scratchpad are logged, providing insight into the user's thought process and estimation strategy.](Estimation_numerical_screenshot_crop.png){#fig-estimation-calc}

@@

Developing a graphics experiment is often a highly iterative process, but it can help to approach the design process by first optimizing the model and data generation method (@sec-model-dev) before spending time on optimizing the specific stimuli or customizing the data collection platform (@sec-protocol-dev).
This is important because the model parameters and data generation process inform the experiment structure and thus impact decisions made downstream.

@@

```{r, echo = F, include = T, message = F, warning = F}
#| label: fig-lof-density-curves
#| fig-cap: "Density plot of the lack of fit statistic showing separation of selected difficulty levels: High (obvious curvature), Medium (noticeable curvature), and Low (almost linear). Each density plot is the result of 1000 simulations from a model $y_i = \\alpha\\cdot e^{\\beta\\cdot x_i + \\epsilon_i} + \\theta$, where $\\epsilon \\sim N(0, \\sigma^2)$. $\\alpha$ and $\\theta$ were selected after manipulation of $\\beta$ and $\\sigma$ to ensure that all data generated had similar $y$ ranges so as not to provide visual cues about model differences outside of the plot curvature."
#| out.width: "\\columnwidth"
#| fig.height: 2.5
#| fig.width: 5

@@

![A lineup from Vanderplas & Hofmann [-@clusters]. Panel 10 shows the clustered target data and panel 17 shows the target data with a strong linear relationship; either of these target panels was the expected choice. Unfortunately, panel 16 has only two bounding ellipses shown, which is an unintentional difference that resulted from a faulty method for assigning clusters to null plots; many participants selected this panel instead of one of the target panels.](lineup-missing-ellipse.png){#fig-lineup-problems}

One reason why it is so difficult to generate sampling models for visual explorations is that our visual system is very good at creating groups and identifying differences between them [@zekiFunctionalSpecializationGeneralization2013;@lupyanConceptualGroupingEffect2008].
This ability can interfere with the natural inclination to use the null sampling models that might be used in equivalent numerical tests when running experiments that use visualizations. 
Numerical tests consider one facet of the data; the power of graphical testing is that it allows examination of a battery of hypotheses simultaneously. 
Unfortunately, that makes designing an effective null sampling model much more difficult than in the numerical test case, as the researcher must effectively control for several different hypotheses when generating null data.

We re-ran the experiment using a different clustering method that controlled the number of points in each group. 
Instead of noticing the number of ellipses, participants instead used the differences in size and shape of the ellipses formed when clustering after the data generating procedure. 
That is, participants could still detect the artificial nature of the induced clusters using other features. 
While it can be difficult to get the data generating method right, it is essential to conducting visual experiments that generalize well beyond the effects shown in a single data set or phenomenon.
This is also why it is critical to include independent replications of the simulated parameters, so that the results reflect variability due to the data generation process, not just the specifics of a single simulated dataset.
The time and effort invested in this step at the outset of the experiment pays dividends when it allows for clear generalization of the experimental results to an entire statistical concept rather than a single data set.

@@

Once the pilot data has been collected and the planned analysis method has been developed, it may be beneficial to pre-register the main study and proposed analysis. 
This can be an effective way to separate exploratory research from confirmatory research involving hypothesis testing. 
However, graphics studies (and other human subjects research) often produce unexpected results that would not be imagined during the pre-registration process (for an example, see @clusters); using only the pre-registered analysis method may disincentivise analyzing (or collecting) the rich, messy open-ended response data which can be incredibly useful but is hard to model and quantify.

@@

```{r, message=FALSE, warning=FALSE, echo = F}
#| label: fig-density-rug
#| fig-cap: "Density of participant estimates for the year (x-value) in which the population reaches 4000 (y-value). Colors are associated to scale - linear (blue) and log (orange) - and vertical lines indicate the true value based on the underlying model equation (black solid) and closest point value based on the simulated data set (black dashed). A jittered rug plot along the $x$-axis shows where participant estimates were made. The plot shows anchoring occurred to the closest point as shown by an increase in density around the dashed line. Density peaks occurred at whole values indicating rounding errors."
#| fig-width: 6
#| fig-height: 3
#| out-width: "\\columnwidth"

@@

# Author Statement
The authors have no conflicts of interest or additional funding to declare in this manuscript.

>>>>

![Direct adjustment of a plot in a perceptual task. In this experiment, designed to assess the strength of the sine illusion, the user adjusts the plot using - and + buttons, which control the strength of a transformation designed to correct the effect of the sine illusion. When the user is satisfied that the lines are of equal length, they select the 'Finished' button to move to the next task. The experiment used a psychophysics experimental design, the method of adjustment, but leveraged the interactive Shiny interface to record the entire sequence of adjustments made by the user for each trial. A demo version of this application can be found at https://shiny.srvanderplas.com/sine-illusion/.  ](images/sine_illusion_screenshot.png){#fig-sine-illusion}

@@

![This question asks users to write out a description of how the population of Ewoks changes over time, without any further cues, to determine whether participants default to multiplicative or additive language descriptions.](images/Estimation_describe_plot_crop.png){#fig-estimation-describe}

@@

![This question asks participants for a numerical estimate, but provides a basic calculator and scratchpad. All user interactions with the calculator and scratchpad are logged, providing insight into the user's thought process and estimation strategy.](images/Estimation_numerical_screenshot_crop.png){#fig-estimation-calc}

@@

Developing a graphics experiment is often a highly iterative process, but it can help to approach the design process by first optimizing the model and data generation method before spending time on optimizing the specific stimuli or customizing the data collection platform.
This is important because the model parameters and data generation process inform the experiment structure and thus impact decisions made downstream.

@@

```{r lof-density-curves, echo = F, include = T, message = F, warning = F}
#| label: fig-lof-density-curves
#| fig-cap: "Density plot of the lack of fit statistic showing separation of selected difficulty levels: High (obvious curvature), Medium (noticeable curvature), and Low (almost linear). Each density plot is the result of 1000 simulations from a model $y_i = \\alpha\\cdot e^{\\beta\\cdot x_i + \\epsilon_i} + \\theta$, where $\\epsilon \\sim N(0, \\sigma^2)$. $\\alpha$ and $\\theta$ were selected after manipulation of $\\beta$ and $\\sigma$ to ensure that all data generated had similar $y$ ranges so as not to provide visual cues about model differences outside of the plot curvature."
#| out.width: "\\columnwidth"
#| fig.height: 2.5
#| fig.width: 5

@@

![A lineup from Vanderplas & Hofmann [-@clusters]. Panel 10 shows the clustered target data and panel 17 shows the target data with a strong linear relationship; either of these target panels was the expected choice. Unfortunately, panel 16 has only two bounding ellipses shown, which is an unintentional difference that resulted from a faulty method for assigning clusters to null plots; many participants selected this panel instead of one of the target panels.](images/lineup-missing-ellipse.png){#fig-lineup-problems}

One reason why it is so difficult to generate sampling models for visual explorations is that our visual system is optimized for identifying differences between groups.
This ability can interfere with the natural  to use the null sampling models that might be used in equivalent numerical tests when running experiments that use visualizations. 
We re-ran the experiment using a different clustering method that controlled the number of points in each group. 
Instead of noticing the number of ellipses, participants instead used the differences in size and shape of the ellipses formed when clustering after the data generating procedure. 
That is, participants could still detect the artificial nature of the induced clusters using other features. 
While it can be difficult to get the data generating method right, it is essential to conducting visual experiments that generalize well beyond the effects shown in a single data set or phenomenon.
This is also why it is critical to include independent replications of the simulated parameters, so that the results reflect variability due to the data generation process, not just the specifics of a single simulated dataset.
The time and effort invested in this step at the outset of the experiment pays dividends when it allows for clear generalization of the experimental results to an entire statistical concept rather than a single data set.

@@

@@

```{r, qe2-density-plot, message=FALSE, warning=FALSE, echo = F}
#| label: fig-density-rug
#| fig-cap: "Density of participant estimates for the year (x-value) in which the population reaches 4000 (y-value). Colors are associated to scale - linear (blue) and log (orange) - and vertical lines indicate the true value based on the underlying model equation (black solid) and closest point value based on the simulated data set (black dashed). A jittered rug plot along the $x$-axis shows where participant estimates were made. The plot shows anchoring occurred to the closest point as shown by an increase in density around the dashed line. Density peaks occurred at whole values indicating rounding errors."
#| fig-width: 6
#| fig-height: 3
#| out-width: "\\columnwidth"

@@

# Author Statements

**Conflict of Interest** The authors have no conflicts of interest to declare in this manuscript.

**Funding** No outside funding was used to create this manuscript.

**Data Availability** This manuscript does not present any newly collected data. Simulated data and code to create many of the images in this paper can be found at https://github.com/srvanderplas/Guide-Testing-Statistical-Graphics.

### 2025-06-04 index-revised.tex
Revision updates

While there are heuristics, general guidelines, and best practices
(Allen \& Erhardt, 2016; Few, 2006; Haemer, 1948; {``Joint Committee on
Standards for Graphic Presentation,''} 1915; Kosslyn, 2006) for creating
useful and visually attractive data displays, the best way to establish
the efficacy of various design decisions is to test the visualization on
humans, evaluating different variants under controlled conditions
(Cleveland et al., 1988; Cleveland \& McGill, 1985). Empirical
assessments of visualizations, when carefully designed, allow
statisticians to determine which representation of the same data is most
effective along one or more dimension(s) of interest: estimation or
prediction accuracy, within or between group comparisons, response time,
and ability to draw real-world decisions are common goals for charts.

@@

\pandocbounded{\includegraphics[keepaspectratio]{sine_illusion_screenshot.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{Estimation_describe_plot_crop.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{Estimation_numerical_screenshot_crop.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-pdf/fig-lof-density-curves-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{lineup-missing-ellipse.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-pdf/fig-density-rug-1.pdf}}

@@

\section{Author Statement}\label{author-statement}

The authors have no conflicts of interest or additional funding to
declare in this manuscript.

@@

\bibitem[\citeproctext]{ref-allenVisualizingScientificData2016}
Allen, E. A., \& Erhardt, E. B. (2016). Visualizing {Scientific} {Data}.
In J. T. Cacioppo, L. G. Tassinary, \& G. G. Berntson (Eds.),
\emph{Handbook of {Psychophysiology}} (4th ed., pp. 679--697). Cambridge
University Press.
https://doi.org/\url{https://doi.org/10.1017/9781107415782.031}

@@

\bibitem[\citeproctext]{ref-d3}
Bostock, M., Ogievetsky, V., \& Heer, J. (2011). D3 {Data}-{Driven}
{Documents}. \emph{IEEE Transactions on Visualization and Computer
Graphics}, \emph{17}(12), 2301--2309.
https://doi.org/\url{https://doi.org/10/b7bhhf}

@@

\bibitem[\citeproctext]{ref-chowdhury2018}
Chowdhury, N. R., Cook, D., Hofmann, H., \& Majumder, M. (2018).
Measuring Lineup Difficulty By Matching Distance Metrics With Subject
Choices in Crowd-Sourced Data. \emph{Journal of Computational and
Graphical Statistics}, \emph{27}(1), 132--145.
https://doi.org/\url{https://doi.org/10.1080/10618600.2017.1356323}

@@

\bibitem[\citeproctext]{ref-clevelandShapeParameterTwoVariable1988}
Cleveland, W. S., McGill, M. E., \& McGill, R. (1988). The {Shape}
{Parameter} of a {Two}-{Variable} {Graph}. \emph{Journal of the American
Statistical Association}, \emph{83}(402), 289--300.
https://doi.org/\url{https://doi.org/10.1080/01621459.1988.10478598}

@@

\bibitem[\citeproctext]{ref-clevelandGraphicalPerceptionGraphical1985}
Cleveland, W. S., \& McGill, R. (1985). Graphical {Perception} and
{Graphical} {Methods} for {Analyzing} {Scientific} {Data}.
\emph{Science}, \emph{229}(4716), 828--833.
https://doi.org/\url{https://doi.org/10.1126/science.229.4716.828}

\bibitem[\citeproctext]{ref-cookFoundationAvailableThinking2021}
Cook, D., Reid, N., \& Tanaka, E. (2021). The {Foundation} {Is}
{Available} for {Thinking} {About} {Data} {Visualization}
{Inferentially}. \emph{Harvard Data Science Review}, \emph{3}(3).
https://doi.org/\url{https://doi.org/10.1162/99608f92.8453435d}

@@

\bibitem[\citeproctext]{ref-croxtonBarChartsCircle1927}
Croxton, F. E., \& Stryker, R. E. (1927). Bar {Charts} {Versus} {Circle}
{Diagrams}. \emph{Journal of the American Statistical Association},
\emph{22}(160), 473--482.
https://doi.org/\url{https://doi.org/10.2307/2276829}

@@

\bibitem[\citeproctext]{ref-desnoyersTaxonomyVisualsScience2011}
Desnoyers, L. (2011). \emph{Toward a {Taxonomy} of {Visuals} in
{Science} {Communication}}. \emph{58}(2), 16.

@@

\bibitem[\citeproctext]{ref-eellsRelativeMeritsCircles1926}
Eells, W. C. (1926). The {Relative} {Merits} of {Circles} and {Bars} for
{Representing} {Component} {Parts}. \emph{Journal of the American
Statistical Association}, \emph{21}(154), 119--132.
https://doi.org/\url{https://doi.org/10.2307/2277140}

@@

\bibitem[\citeproctext]{ref-thinkaloud}
Haak, M., De Jong, M., \& Schellens, P. (2003). Retrospective vs.
Concurrent think-aloud protocols: Testing the usability of an online
library catalogue. \emph{Behaviour \& {IT}}, \emph{22}, 339--351.
\url{https://doi.org/10.1080/0044929031000}

\bibitem[\citeproctext]{ref-haemerDoubleScalesAre1948}
Haemer, K. W. (1948). Double {Scales} are {Dangerous}. \emph{The
American Statistician}, \emph{2}(3), 24--24.
https://doi.org/\url{https://doi.org/10.1080/00031305.1948.10501588}

\bibitem[\citeproctext]{ref-haiderStrategiesDetectingDifference2021}
Haider, J. D., Pohl, M., Beecham, R., \& Dykes, J. (2021). Strategies
for detecting difference in~map line-up tasks. In C. Ardito, R.
Lanzilotti, A. Malizia, H. Petrie, A. Piccinno, G. Desolda, \& K. Inkpen
(Eds.), \emph{Human-computer interaction -- {INTERACT} 2021} (pp.
558--578). Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-85613-7_36}

@@

\bibitem[\citeproctext]{ref-heitjanIgnorabilityCoarseData1991}
Heitjan, D. F., \& Rubin, D. B. (1991). Ignorability and {Coarse}
{Data}. \emph{The Annals of Statistics}, \emph{19}(4), 2244--2253.
https://doi.org/\url{https://doi.org/10.1214/aos/1176348396}

@@

\bibitem[\citeproctext]{ref-vonhuhnFurtherStudiesGraphic1927}
Huhn, R. von. (1927). Further {Studies} in the {Graphic} {Use} of
{Circles} and {Bars}: {A} {Discussion} of the {Eell}'s {Experiment}.
\emph{Journal of the American Statistical Association}, \emph{22}(157),
31--39. https://doi.org/\url{https://doi.org/10.2307/2277346}

\bibitem[\citeproctext]{ref-hullmanDesigningInteractiveExploratory2021}
Hullman, J., \& Gelman, A. (2021). Designing for {Interactive}
{Exploratory} {Data} {Analysis} {Requires} {Theories} of {Graphical}
{Inference}. \emph{Harvard Data Science Review}, \emph{3}(3).
https://doi.org/\url{https://doi.org/10.1162/99608f92.3ab8a587}

\bibitem[\citeproctext]{ref-asme-standards-graphics}
Joint committee on standards for graphic presentation. (1915).
\emph{Publications of the American Statistical Association},
\emph{14}(112), 790--797. \url{http://www.jstor.org/stable/2965153}

@@

\bibitem[\citeproctext]{ref-kirschenbaum2003comparative}
Kirschenbaum, S. S. (2003). Comparative {Cognitive} {Task} {Analysis}:
{The} {Cognition} of {Weather} {Forecasting}. \emph{Proceedings of the
{Human} {Factors} and {Ergonomics} {Society} {Annual} {Meeting}},
\emph{47}, 473--477.
https://doi.org/\url{https://doi.org/10.1177/154193120304700347}

@@

\bibitem[\citeproctext]{ref-loyDiagnosticToolsHierarchical2013}
Loy, A., \& Hofmann, H. (2013). Diagnostic tools for hierarchical linear
models. \emph{Wiley Interdisciplinary Reviews: Computational
Statistics}, \emph{5}(1), 48--61.
https://doi.org/\url{https://doi.org/10.1002/wics.1238}

\bibitem[\citeproctext]{ref-luModelingJustNoticeable2022}
Lu, M., Lanir, J., Wang, C., Yao, Y., Zhang, W., Deussen, O., \& Huang,
H. (2022). Modeling {Just} {Noticeable} {Differences} in {Charts}.
\emph{IEEE Transactions on Visualization and Computer Graphics},
\emph{28}(1), 718--726.
https://doi.org/\url{https://doi.org/10.1109/TVCG.2021.3114874}

\bibitem[\citeproctext]{ref-majumderValidationVisualStatistical2013}
Majumder, M., Hofmann, H., \& Cook, D. (2013). Validation of visual
statistical inference, applied to linear models. \emph{Journal of the
American Statistical Association}, \emph{108}(503), 942--956.
https://doi.org/\url{https://doi.org/10/f5gntt}

@@

@@

\bibitem[\citeproctext]{ref-robinson2023you}
Robinson, E. A., Howard, R., \& VanderPlas, S. (2023b). {``You draw
it''}: Implementation of visually fitted trends with r2d3. \emph{Journal
of Data Science}, \emph{21}(2), 281--294.

@@

\bibitem[\citeproctext]{ref-tourangeau_rips_rasinski_2000}
Tourangeau, R., Rips, L. J., \& Rasinski, K. (2000). \emph{The
psychology of survey response}. Cambridge University Press.
https://doi.org/\url{https://doi.org/10.1017/CBO9780511819322}

\bibitem[\citeproctext]{ref-traftonTurningPicturesNumbers2000}
Trafton, G. J., Kirschenbaum, S. S., Tsui, T. L., Miyamoto, R. T.,
Ballas, J. A., \& Raymond, P. D. (2000). Turning pictures into numbers:
Extracting and generating information from complex visualizations.
\emph{International Journal of Human-Computer Studies}, \emph{53}(5),
827--850. \url{https://doi.org/10.1006/ijhc.2000.0419}

@@

\bibitem[\citeproctext]{ref-ushakovStatisticalAnalysisRounded2017}
Ushakov, N. G., \& Ushakov, V. G. (2017). Statistical analysis of
rounded data: {Recovering} of information lost due to rounding.
\emph{Journal of the Korean Statistical Society}, \emph{46}(3),
426--437.
https://doi.org/\url{https://doi.org/10.1016/j.jkss.2017.01.003}

\bibitem[\citeproctext]{ref-valentinDesigningOptimalBehavioral2024}
Valentin, S., Kleinegesse, S., Bramley, N. R., Seriès, P., Gutmann, M.
U., \& Lucas, C. G. (2024). Designing optimal behavioral experiments
using machine learning. \emph{{eLife}}, \emph{13}.
\url{https://doi.org/10.7554/eLife.86224}

\bibitem[\citeproctext]{ref-vanderplasDesigningGraphicsRequires2021}
Vanderplas, S. (2021). Designing {Graphics} {Requires} {Useful}
{Experimental} {Testing} {Frameworks} and {Graphics} {Derived} {From}
{Empirical} {Results}. \emph{Harvard Data Science Review}, \emph{3}(3).
https://doi.org/\url{https://doi.org/10.1162/99608f92.7d099fd0}

@@

\bibitem[\citeproctext]{ref-vanderplasTestingStatisticalCharts2020}
Vanderplas, S., Cook, D., \& Hofmann, H. (2020). Testing {Statistical}
{Charts}: {What} {Makes} a {Good} {Graph}? \emph{Annual Review of
Statistics and Its Application}, \emph{7}(1).
https://doi.org/\url{https://doi.org/10.1146/annurev-statistics-031219-041252}

\bibitem[\citeproctext]{ref-vanderplasFramedReproducingRevisiting2019}
Vanderplas, S., Goluch, R., \& Hofmann, H. (2019). Framed! {Reproducing}
and {Revisiting} 150 year old charts. \emph{Journal of Computational and
Graphical Statistics}.
https://doi.org/\url{https://doi.org/10.1080/10618600.2018.1562937}

\bibitem[\citeproctext]{ref-sineillusion}
Vanderplas, S., \& Hofmann, H. (2015). Signs of the {Sine}
{Illusion}---{Why} {We} {Need} to {Care}. \emph{Journal of Computational
and Graphical Statistics}, \emph{24}(4), 1170--1190.
https://doi.org/\url{https://doi.org/10.1080/10618600.2014.951547}

@@

\bibitem[\citeproctext]{ref-vanderplasStatisticalSignificanceCalculations2021}
Vanderplas, S., Röttger, C., Cook, D., \& Hofmann, H. (2021).
Statistical significance calculations for scenarios in visual inference.
\emph{Stat}, \emph{10}(1).
https://doi.org/\url{https://doi.org/10.1002/sta4.337}

@@

\bibitem[\citeproctext]{ref-wickhamGraphicalInferenceInfovis2010}
Wickham, H., Cook, D., Hofmann, H., \& Buja, A. (2010). Graphical
inference for infovis. \emph{IEEE Transactions on Visualization and
Computer Graphics}, \emph{16}(6), 973--979.
https://doi.org/\url{https://doi.org/10.1109/TVCG.2010.161}

>>>>

While there are heuristics, general guidelines, and best practices
(Allen \& Erhardt, 2016; Few, 2006; Haemer, 1948; Kosslyn, 2006;
presentation, 1915) for creating useful and visually attractive data
displays, the best way to establish the efficacy of various design
decisions is to test the visualization on humans, evaluating different
variants under controlled conditions (Cleveland et al., 1988; Cleveland
\& McGill, 1985). Empirical assessments of visualizations, when
carefully designed, allow statisticians to determine which
representation of the same data is most effective along one or more
dimension(s) of interest: estimation or prediction accuracy, within or
between group comparisons, response time, and ability to draw real-world
decisions are common goals for charts.

@@

\pandocbounded{\includegraphics[keepaspectratio]{images/sine_illusion_screenshot.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{images/Estimation_describe_plot_crop.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{images/Estimation_numerical_screenshot_crop.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index-revised_files/figure-pdf/fig-lof-density-curves-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{images/lineup-missing-ellipse.png}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index-revised_files/figure-pdf/fig-density-rug-1.pdf}}

@@

\section{Author Statements}\label{author-statements}

\textbf{Conflict of Interest} The authors have no conflicts of interest
to declare in this manuscript.

\textbf{Funding} No outside funding was used to create this manuscript.

\textbf{Data Availability} This manuscript does not present any newly
collected data. Simulated data and code to create many of the images in
this paper can be found at
https://github.com/srvanderplas/Guide-Testing-Statistical-Graphics.

@@

\bibitem[\citeproctext]{ref-allenVisualizingScientificData2016}
Allen, E. A., \& Erhardt, E. B. (2016). Visualizing {Scientific} {Data}.
In J. T. Cacioppo, L. G. Tassinary, \& G. G. Berntson (Eds.),
\emph{Handbook of {Psychophysiology}} (4th ed., pp. 679--697). Cambridge
University Press. \url{https://doi.org/10.1017/9781107415782.031}

@@

\bibitem[\citeproctext]{ref-d3}
Bostock, M., Ogievetsky, V., \& Heer, J. (2011). D3 {Data}-{Driven}
{Documents}. \emph{IEEE Transactions on Visualization and Computer
Graphics}, \emph{17}(12), 2301--2309. \url{https://doi.org/10/b7bhhf}

@@

\bibitem[\citeproctext]{ref-chowdhury2018}
Chowdhury, N. R., Cook, D., Hofmann, H., \& Majumder, M. (2018).
Measuring Lineup Difficulty By Matching Distance Metrics With Subject
Choices in Crowd-Sourced Data. \emph{Journal of Computational and
Graphical Statistics}, \emph{27}(1), 132--145.
\url{https://doi.org/10.1080/10618600.2017.1356323}

@@

\bibitem[\citeproctext]{ref-clevelandShapeParameterTwoVariable1988}
Cleveland, W. S., McGill, M. E., \& McGill, R. (1988). The {Shape}
{Parameter} of a {Two}-{Variable} {Graph}. \emph{Journal of the American
Statistical Association}, \emph{83}(402), 289--300.
\url{https://doi.org/10.1080/01621459.1988.10478598}

@@

\bibitem[\citeproctext]{ref-clevelandGraphicalPerceptionGraphical1985}
Cleveland, W. S., \& McGill, R. (1985). Graphical {Perception} and
{Graphical} {Methods} for {Analyzing} {Scientific} {Data}.
\emph{Science}, \emph{229}(4716), 828--833.
\url{https://doi.org/10.1126/science.229.4716.828}

\bibitem[\citeproctext]{ref-cookFoundationAvailableThinking2021}
Cook, D., Reid, N., \& Tanaka, E. (2021). The {Foundation} {Is}
{Available} for {Thinking} {About} {Data} {Visualization}
{Inferentially}. \emph{Harvard Data Science Review}, \emph{3}(3).
\url{https://doi.org/10.1162/99608f92.8453435d}

@@

\bibitem[\citeproctext]{ref-croxtonBarChartsCircle1927}
Croxton, F. E., \& Stryker, R. E. (1927). Bar {Charts} {Versus} {Circle}
{Diagrams}. \emph{Journal of the American Statistical Association},
\emph{22}(160), 473--482. \url{https://doi.org/10.2307/2276829}

@@

\bibitem[\citeproctext]{ref-desnoyersTaxonomyVisualsScience2011}
Desnoyers, L. (2011). Toward a {Taxonomy} of {Visuals} in {Science}
{Communication}. \emph{Technical Communication}, \emph{58}(2), 16.

@@

\bibitem[\citeproctext]{ref-eellsRelativeMeritsCircles1926}
Eells, W. C. (1926). The {Relative} {Merits} of {Circles} and {Bars} for
{Representing} {Component} {Parts}. \emph{Journal of the American
Statistical Association}, \emph{21}(154), 119--132.
\url{https://doi.org/10.2307/2277140}

@@

\bibitem[\citeproctext]{ref-thinkaloud}
Haak, M., De Jong, M., \& Schellens, P. (2003). Retrospective vs.
Concurrent think-aloud protocols: Testing the usability of an online
library catalogue. \emph{Behaviour \& {Information Technology}},
\emph{22}, 339--351. \url{https://doi.org/10.1080/0044929031000}

\bibitem[\citeproctext]{ref-haemerDoubleScalesAre1948}
Haemer, K. W. (1948). Double {Scales} are {Dangerous}. \emph{The
American Statistician}, \emph{2}(3), 24--24.
\url{https://doi.org/10.1080/00031305.1948.10501588}

\bibitem[\citeproctext]{ref-haiderStrategiesDetectingDifference2021}
Haider, J. D., Pohl, M., Beecham, R., \& Dykes, J. (2021). Strategies
for detecting difference in map line-up tasks. In C. Ardito, R.
Lanzilotti, A. Malizia, H. Petrie, A. Piccinno, G. Desolda, \& K. Inkpen
(Eds.), \emph{Human-computer interaction - {INTERACT} 2021} (pp.
558--578). Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-85613-7_36}

@@

\bibitem[\citeproctext]{ref-heitjanIgnorabilityCoarseData1991}
Heitjan, D. F., \& Rubin, D. B. (1991). Ignorability and {Coarse}
{Data}. \emph{The Annals of Statistics}, \emph{19}(4), 2244--2253.
\url{https://doi.org/10.1214/aos/1176348396}

@@

\bibitem[\citeproctext]{ref-vonhuhnFurtherStudiesGraphic1927}
Huhn, R. von. (1927). Further {Studies} in the {Graphic} {Use} of
{Circles} and {Bars}: {A} {Discussion} of the {Eell}'s {Experiment}.
\emph{Journal of the American Statistical Association}, \emph{22}(157),
31--39. \url{https://doi.org/10.2307/2277346}

\bibitem[\citeproctext]{ref-hullmanDesigningInteractiveExploratory2021}
Hullman, J., \& Gelman, A. (2021). Designing for {Interactive}
{Exploratory} {Data} {Analysis} {Requires} {Theories} of {Graphical}
{Inference}. \emph{Harvard Data Science Review}, \emph{3}(3).
\url{https://doi.org/10.1162/99608f92.3ab8a587}

@@

\bibitem[\citeproctext]{ref-kirschenbaum2003comparative}
Kirschenbaum, S. S. (2003). Comparative {Cognitive} {Task} {Analysis}:
{The} {Cognition} of {Weather} {Forecasting}. \emph{Proceedings of the
{Human} {Factors} and {Ergonomics} {Society} {Annual} {Meeting}},
\emph{47}, 473--477. \url{https://doi.org/10.1177/154193120304700347}

@@

\bibitem[\citeproctext]{ref-loyDiagnosticToolsHierarchical2013}
Loy, A., \& Hofmann, H. (2013). Diagnostic tools for hierarchical linear
models. \emph{Wiley Interdisciplinary Reviews: Computational
Statistics}, \emph{5}(1), 48--61.
\url{https://doi.org/10.1002/wics.1238}

\bibitem[\citeproctext]{ref-luModelingJustNoticeable2022}
Lu, M., Lanir, J., Wang, C., Yao, Y., Zhang, W., Deussen, O., \& Huang,
H. (2022). Modeling just noticeable differences in charts. \emph{{IEEE}
Transactions on Visualization and Computer Graphics}, \emph{28}(1),
718--726. \url{https://doi.org/10.1109/TVCG.2021.3114874}

\bibitem[\citeproctext]{ref-majumderValidationVisualStatistical2013}
Majumder, M., Hofmann, H., \& Cook, D. (2013). Validation of visual
statistical inference, applied to linear models. \emph{Journal of the
American Statistical Association}, \emph{108}(503), 942--956.
\url{https://doi.org/10/f5gntt}

@@

\bibitem[\citeproctext]{ref-asme-standards-graphics}
presentation, J. committee on standards for graphic. (1915). Preliminary
report published for the purpose of inviting suggestions for the benefit
of the committee. \emph{Publications of the American Statistical
Association}, \emph{14}(112), 790--797.
\url{http://www.jstor.org/stable/2965153}

@@

\bibitem[\citeproctext]{ref-robinson2023you}
Robinson, E. A., Howard, R., \& VanderPlas, S. (2023b). 'You draw it':
Implementation of visually fitted trends with r2d3. \emph{Journal of
Data Science}, \emph{21}(2), 281--294.

@@

\bibitem[\citeproctext]{ref-tourangeau_rips_rasinski_2000}
Tourangeau, R., Rips, L. J., \& Rasinski, K. (2000). \emph{The
psychology of survey response}. Cambridge University Press.
\url{https://doi.org/10.1017/CBO9780511819322}

\bibitem[\citeproctext]{ref-traftonTurningPicturesNumbers2000}
Trafton, G. J., Kirschenbaum, S. S., Tsui, T. L., Miyamoto, R. T.,
Ballas, J. A., \& Raymond, P. D. (2000). Turning pictures into numbers:
Extracting and generating information from complex visualizations.
\emph{International Journal of Human-Computer Studies}, \emph{53}(5),
827--850. \href{https://10.1006/ijhc.2000.0419}{10.1006/ijhc.2000.0419}

@@

\bibitem[\citeproctext]{ref-ushakovStatisticalAnalysisRounded2017}
Ushakov, N. G., \& Ushakov, V. G. (2017). Statistical analysis of
rounded data: {Recovering} of information lost due to rounding.
\emph{Journal of the Korean Statistical Society}, \emph{46}(3),
426--437. \url{https://doi.org/10.1016/j.jkss.2017.01.003}

\bibitem[\citeproctext]{ref-valentinDesigningOptimalBehavioral2024}
Valentin, S., Kleinegesse, S., Bramley, N. R., Seriés, P., Gutmann, M.
U., \& Lucas, C. G. (2024). Designing optimal behavioral experiments
using machine learning. \emph{{eLife}}, \emph{13}.
\url{https://doi.org/10.7554/eLife.86224}

\bibitem[\citeproctext]{ref-vanderplasDesigningGraphicsRequires2021}
Vanderplas, S. (2021). Designing {Graphics} {Requires} {Useful}
{Experimental} {Testing} {Frameworks} and {Graphics} {Derived} {From}
{Empirical} {Results}. \emph{Harvard Data Science Review}, \emph{3}(3).
\url{https://doi.org/10.1162/99608f92.7d099fd0}

@@

\bibitem[\citeproctext]{ref-vanderplasTestingStatisticalCharts2020}
Vanderplas, S., Cook, D., \& Hofmann, H. (2020). Testing {Statistical}
{Charts}: {What} {Makes} a {Good} {Graph}? \emph{Annual Review of
Statistics and Its Application}, \emph{7}(1).
\url{https://doi.org/10.1146/annurev-statistics-031219-041252}

\bibitem[\citeproctext]{ref-vanderplasFramedReproducingRevisiting2019}
Vanderplas, S., Goluch, R., \& Hofmann, H. (2019). Framed! {Reproducing}
and {Revisiting} 150 year old charts. \emph{Journal of Computational and
Graphical Statistics}.
\url{https://doi.org/10.1080/10618600.2018.1562937}

\bibitem[\citeproctext]{ref-sineillusion}
Vanderplas, S., \& Hofmann, H. (2015). Signs of the {Sine} {Illusion} --
{Why} {We} {Need} to {Care}. \emph{Journal of Computational and
Graphical Statistics}, \emph{24}(4), 1170--1190.
\url{https://doi.org/10.1080/10618600.2014.951547}

@@

\bibitem[\citeproctext]{ref-vanderplasStatisticalSignificanceCalculations2021}
Vanderplas, S., Röttger, C., Cook, D., \& Hofmann, H. (2021).
Statistical significance calculations for scenarios in visual inference.
\emph{Stat}, \emph{10}(1). \url{https://doi.org/10.1002/sta4.337}

@@

\bibitem[\citeproctext]{ref-wickhamGraphicalInferenceInfovis2010}
Wickham, H., Cook, D., Hofmann, H., \& Buja, A. (2010). Graphical
inference for infovis. \emph{IEEE Transactions on Visualization and
Computer Graphics}, \emph{16}(6), 973--979.
\url{https://doi.org/10.1109/TVCG.2010.161}
